---
title: "Initial Korean covid 19"
author: "Steven Garcia"
date: "11/1/2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### These data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well 

```{r}
str(kRdata)
```



  
# Your Question - Is dying independent of sex?

$H_0$: There is no difference between the connection of COVID-19 deaths and the death based on sex. 

$H_A$: There is a difference between the connection of COVID-19 deaths and the death associated based on sex. 

## Method

The respone variable is categorical and the explanitory variable is categorical. We will be utilizing the helper file Cat~Cat. 


## Graphical results - use ggplot commands to do this  


```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
According to the graph the plot indicates that more men die compared to women who contract COVID-19.
  
### Numerical Summaries  

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
```
Among those who died, almost twice as many men dies when compared to women. 
  


## Inferential Result 

## Inferential results 

```{r}
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
```

The chart indicates that fisher exact test has a P-value is 0.00000004, which is less than 0.05, so we reject the null hypothesis. The samle indicates that males are 2.3 times more likely to die from COVID-19 in Korea.

## Conclusions

In conclusion, we see that there is no difference between the effect of COVID-19 and the death associated based on sex. The P-value is less than 0.05, so we rejected the null hypothesis. The samle indicates that males are 2.3 times more likely to die from COVID-19 in Korea.